A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-11-07 latex/main.tex
Updates from Overleaf

\section*{DISCUSSION}
There are two primary contributions of this work. The first is an
ARG for 2.48 million SARS-CoV-2 genomes.
Through multiple layers of validation against results from the 
literature, we have demonstrated that this ARG synthesises
the evolutionary features of genetic inheritance, point mutations,
deletions and recombination in a single cohesive structure.
The ARG is freely available and easily accessible using the 
tskit software ecosystem, a rich and efficient platform for 
recombination-aware analyses widely used in population genomics.
This combination of a deeply validated pandemic-scale ARG with 
a mature and highly capable software platform forms a unique resource.
Previous landmark ARGs~\cite{Wohns2022,Anderson2023genes}
have helped spur the development of ARG-aware methods in human 
genomics~\cite{Mejia2025using,Grundler2025geographic,
Lehmann2025args,Terhorst2025accelerated}.
We hope that the ARG provided here will similarly
enable a range of recombination-aware analyses of SARS-CoV-2, and 
catalyse a wave of methodological developments integrating
recombination into pathogen genomics.
We envisage the ARG as a long-term stable resource---building on the 
high-quality, freely available reassemblies of
the Viridian dataset~\cite{Hunt2024}---to facilitate the study 
of the pandemic phase of SARS-CoV-2 in an open and reproducible manner.
% NOTE: still missing the "What" bit suggested by Jotun, where we 
% list out some of the stuff that could be done. I'm not convinced
% this works, though.

>>>>

\section*{DISCUSSION}
There are two primary contributions of this work. The first is an
ARG for 2.48 million SARS-CoV-2 genomes.
Through multiple layers of validation against results from the 
literature, we have demonstrated that this ARG synthesises
the evolutionary features of genetic inheritance, point mutations,
deletions and recombination in a single cohesive structure.
The ARG is freely available and easily accessible using the 
tskit software ecosystem, a rich and efficient platform for 
recombination-aware analyses widely used in population genomics.
This combination of a deeply validated pandemic-scale ARG with 
a mature and highly capable software platform forms a unique resource.
Previous landmark ARGs~\cite{Wohns2022,Anderson2023genes}
have helped spur the development of ARG-aware methods in human 
genomics~\cite{Mejia2025using,Grundler2025geographic,
Lehmann2025args,Terhorst2025accelerated}.
We hope that the ARG provided here will similarly
enable a range of recombination-aware analyses of SARS-CoV-2, and 
catalyse a wave of methodological developments integrating
recombination into pathogen genomics.
For instance, annotating the ARG with experimentally derived
fitness estimates for mutations can help elucidate selection
pressures on difference variants and understand their function;
analysing the genomic positions of mutations occurring on long
branches can illuminate the evolution of saltational variants.
We envisage the ARG as a long-term stable resource---building on the 
high-quality, freely available reassemblies of
the Viridian dataset~\cite{Hunt2024}---to facilitate the study 
of the pandemic phase of SARS-CoV-2 in an open and reproducible manner.
% NOTE: still missing the "What" bit suggested by Jotun, where we 
% list out some of the stuff that could be done. I'm not convinced
% this works, though.
